% Topological Positioning Scripting Language
% A Domain-Specific Language for Evidence-Based Geolocation
%
% Compile with:
%   pdflatex topological-positioning-scripting
%   bibtex topological-positioning-scripting
%   pdflatex topological-positioning-scripting
%   pdflatex topological-positioning-scripting

\documentclass[twocolumn,10pt]{article}

%-----------------------------------------------------------------------
% Packages
%-----------------------------------------------------------------------
\usepackage[margin=1in,columnsep=0.25in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{enumitem}
\usepackage[expansion=false]{microtype}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{abstract}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{cite}
\usepackage{url}
\usepackage{stmaryrd}

%-----------------------------------------------------------------------
% Page style
%-----------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Topological Positioning Scripting Language}}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

%-----------------------------------------------------------------------
% Theorem environments
%-----------------------------------------------------------------------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

%-----------------------------------------------------------------------
% Listings / PoSL syntax
%-----------------------------------------------------------------------
\definecolor{poslkw}{RGB}{0,0,180}
\definecolor{poslcomment}{RGB}{60,128,49}
\definecolor{poslstring}{RGB}{163,21,21}
\definecolor{poslnumber}{RGB}{9,134,88}
\definecolor{codebg}{RGB}{248,248,248}

\lstdefinelanguage{PoSL}{
  morekeywords={query,locate,given,within,motion,evidence,
                proposition,with,confidence,from,requires,
                pattern,registry,measure,confidence_floor,
                collect,support,contradict,scale,depth,
                return,as,synchronized,phase,compile,execute},
  sensitive=true,
  morecomment=[l]{\#},
  morecomment=[s]{(*}{*)},
  morestring=[b]",
  keywordstyle=\color{poslkw}\bfseries,
  commentstyle=\color{poslcomment}\itshape,
  stringstyle=\color{poslstring},
  numberstyle=\tiny\color{poslnumber},
  basicstyle=\footnotesize\ttfamily,
  backgroundcolor=\color{codebg},
  frame=single,
  framesep=4pt,
  xleftmargin=8pt,
  xrightmargin=4pt,
  breaklines=true,
  showstringspaces=false,
  tabsize=2,
  captionpos=b
}

\lstset{language=PoSL}

%-----------------------------------------------------------------------
% Section formatting
%-----------------------------------------------------------------------
\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}

%-----------------------------------------------------------------------
% Math macros
%-----------------------------------------------------------------------
\newcommand{\calC}{\mathcal{C}}     % cell/confidence
\newcommand{\calP}{\mathcal{P}}     % partition/proposition
\newcommand{\calE}{\mathcal{E}}     % evidence
\newcommand{\calM}{\mathcal{M}}     % motion
\newcommand{\calS}{\mathcal{S}}     % state space
\newcommand{\calT}{\mathcal{T}}     % trajectory
\newcommand{\llb}{\llbracket}
\newcommand{\rrb}{\rrbracket}
\newcommand{\dcomp}{\;\|\;}         % parallel composition
\DeclareMathOperator{\Leb}{Leb}
\DeclareMathOperator{\depth}{depth}
\DeclareMathOperator{\conf}{conf}
\DeclareMathOperator{\traj}{traj}
\DeclareMathOperator{\cell}{cell}
\DeclareMathOperator{\evidence}{evidence}

%-----------------------------------------------------------------------
% Title
%-----------------------------------------------------------------------
\title{%
  \LARGE\bfseries Topological Positioning Scripting Language:\\
  Evidence-Based Geolocation Through\\
  Partition State Classification and\\
  Compositional Confidence Metrics%
}

\author{Kundai Farai Sachikonye\\
\small Chair of Proteomics and Bioanalytics, Technical University of Munich\\
\small AIMe Registry for Artificial Intelligence\\
\small \texttt{kundai.sachikonye@wzw.tum.de}}

\date{May 2026}

%=======================================================================
\begin{document}
%=======================================================================

\twocolumn[{%
\maketitle
\begin{@twocolumnfalse}

%-----------------------------------------------------------------------
\begin{abstract}
We introduce \emph{PoSL} (Positioning Scripting Language), a domain-specific
language for expressing geolocation queries as structured propositions supported
by evidence from multiple measurement modalities. Unlike traditional
positioning systems that output raw coordinates, PoSL programs query the
existence of positions consistent with observed evidence and return
positions annotated with rigorous confidence metrics. The core theoretical
contribution is a confidence framework based on \emph{composition inflation}:
for $n$ independent evidence sources distributed over $d$ measurement
channels, the number of categorically distinguishable position hypotheses
grows as $\Gamma(n,d) = d\,(1+d)^{n-1}$. This exponential growth in
hypothesis space directly bounds the confidence of position estimates.
We formalize PoSL with a complete grammar, denotational semantics, and
a compile-execute phase separation that enforces correctness by construction.
Security properties include structural immunity to query injection
(no position parser exists), replay resistance via monotone epoch counters,
and bounded hypothesis surface (finite, enumerable at compile time).
We demonstrate applications in sports analytics, autonomous navigation,
and infrastructure monitoring, with validation on real-world atmospheric
measurement networks.
\end{abstract}
\vspace{1em}

\end{@twocolumnfalse}
}]

%=======================================================================
\section{Introduction}
\label{sec:intro}
%=======================================================================

\subsection{Motivation}

Geolocation systems face a fundamental asymmetry: they must output a single
coordinate pair $(x,y)$ from noisy, multimodal measurements. This design
forces all uncertainty into the output—either the system returns a point
estimate (losing confidence bounds) or returns a bounding box (losing precision).

Consider instead a different question: \emph{Given my observations, what
positions are consistent with the evidence?} This inversion has deep
consequences. Rather than forcing a measurement into a fixed output format,
we classify the measurement into a categorical space and reason over which
positions explain that category.

A sports analytics example: an athlete's position is not intrinsic to pixel
coordinates in a video. Instead, the question is: \emph{What real-world
position would produce these pixel observations, given the camera geometry
and field structure?} Multiple real-world positions might produce similar
pixels (depth ambiguity, occlusion). The evidence is the pixel configuration;
the position hypothesis is what we infer from that evidence.

Traditional approaches (GNSS, trilateration, SLAM) all embed geometric
assumptions. PoSL takes a step back and treats positioning as a hypothesis
selection problem: organize the space of possible positions, characterize
each region by its evidence signature, and use the signature to classify
the measurement.

\subsection{Core Idea: Position as Evidence Classification}

A positioning query in PoSL is a \emph{proposition}:

\begin{quotation}
\textit{``Is the observed state consistent with the hypothesis that
the target is in region $R$ with confidence $C$?''}
\end{quotation}

The answer is not a point $(x,y)$ but a triple: (position region, confidence
score, composition depth). The confidence score is grounded in the
\emph{composition inflation} principle: the more independent evidence sources
that align on a single position region, the exponentially larger the space
of alternative positions that would contradict that evidence.

\subsection{Contributions}

\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item \textbf{PoSL Language.} A complete domain-specific language with
    BNF grammar, type system, and operational semantics for expressing
    positioning queries as evidence-based propositions.

  \item \textbf{Composition Inflation Theorem.} Proof that $\Gamma(n,d) = d\,(1+d)^{n-1}$
    hypothesis alternatives are distinguishable for $n$ evidence sources
    over $d$ channels, providing formal confidence bounds.

  \item \textbf{Phase-Separated Runtime.} COMPILE and EXECUTE phases
    are mutually exclusive states, enforcing correctness by separation
    rather than validation.

  \item \textbf{Structural Security.} No position parser exists; all
    computation is hypothesis classification and evidence aggregation.
    Replay attacks fail under monotone epoch counters. Attack surface
    is finite and enumerable.

  \item \textbf{Evidence Integration Framework.} Formalization of how
    heterogeneous measurements (pixels, atmospheric data, ranging,
    acceleration) are mapped to a unified confidence space.

  \item \textbf{Multi-Domain Applications.} Validation across sports
    analytics, drone positioning, terrain-based geolocation, and
    infrastructure sensing.
\end{enumerate}

%=======================================================================
\section{Background and Related Work}
\label{sec:background}
%=======================================================================

\subsection{Positioning Paradigms}

Classical geolocation systems fall into three categories:

\textbf{GNSS-based:} Trilateration from satellite signals. Strengths:
global coverage, autonomous. Weaknesses: requires clear sky, susceptible
to spoofing, 5--50~m accuracy outdoors, unusable indoors.

\textbf{SLAM-based:} Simultaneous localization and mapping. Build a map
while localizing within it. Strengths: indoor coverage, high precision
(1--10~cm). Weaknesses: requires rich features, computationally expensive,
must be seeded with known position or loop-closure constraints.

\textbf{Trilateration/Multilateration:} Time-of-arrival from beacons or
reflectors. Strengths: can be precise (cm to m scale). Weaknesses: requires
time synchronization, beacon infrastructure, line-of-sight.

PoSL takes a fourth approach: \emph{evidence classification}. The target
position is identified not by measuring distance to known points, but by
classifying the observed evidence signature into the position regions whose
characteristic signatures match.

\subsection{Hypothesis Testing in Positioning}

Bayesian positioning frameworks \cite{thrun2005probabilistic} compute posteriors
over position space by likelihood weighting. PoSL differs in that it
\emph{categorizes} the evidence first (e.g., ``this video segment shows
fast horizontal motion'') and then matches that category to position regions.
This is more robust to measurement noise that would derail likelihood-based
methods: an outlier video frame doesn't destroy the positional hypothesis
if the bulk of the evidence is consistent.

\subsection{Domain-Specific Languages for Spatial Reasoning}

Prior work on spatial DSLs (GeoSPARQL, PostGIS, Spatial4J) focuses on
geometric query and storage. PoSL is complementary: it focuses on
\emph{inferring} position from evidence, not querying a fixed geometry.

%=======================================================================
\section{The Topological Positioning Model}
\label{sec:model}
%=======================================================================

\subsection{Position Space and Partition}

\begin{definition}[Position Space]
\label{def:posspace}
A \emph{position space} is a measurable space $(\Omega, \mathcal{B}, \mu)$
where $\Omega \subset \mathbb{R}^d$ is a bounded region (e.g., a sports field,
building floor, terrain), $\mathcal{B}$ is the Borel $\sigma$-algebra, and
$\mu$ is the Lebesgue measure. A \emph{position partition} is a finite
collection $\calP = \{\Omega_1, \ldots, \Omega_k\}$ such that:
\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item[(i)]   $\bigcup_{i=1}^{k} \Omega_i = \Omega$ (coverage),
  \item[(ii)]  $\Omega_i \cap \Omega_j = \emptyset$ for $i \neq j$
    (disjointness, up to measure-zero boundaries),
  \item[(iii)] $\mu(\Omega_i) > 0$ for all $i$ (positive measure).
\end{enumerate}
\end{definition}

Each $\Omega_i$ is a \emph{position region}. The partition can be regular
(grid) or irregular (semantic regions like ``left sideline'', ``goal area'').

\subsection{Evidence Modalities and Measurement Channels}

\begin{definition}[Measurement Channel]
\label{def:channel}
A \emph{measurement channel} is a pair $\chi = (s, \calC_\chi)$ where $s$ is
a modality identifier (e.g., ``video'', ``pressure'', ``acoustic'') and
$\calC_\chi$ is a finite partition of the observable space for that modality.
Each cell $c \in \calC_\chi$ represents a measurable equivalence class of
observations (e.g., ``motion blur detected'', ``pressure reading in range
[1005, 1010]~hPa'').
\end{definition}

\begin{definition}[Signature]
\label{def:sig}
For a position region $\Omega_i$ and channel $\chi$, the \emph{signature}
is a probability distribution $\sigma_\chi(i) : \calC_\chi \to [0,1]$
representing the likelihood that an observation falls in each cell
given the target is in $\Omega_i$. Formally:
\begin{equation}
  \sigma_\chi(i)(c) \;=\; \Pr[\text{obs} \in c \mid \text{target} \in \Omega_i].
  \label{eq:sig}
\end{equation}
\end{definition}

The signature is computed offline from reference data, domain knowledge,
or physics models. For video: signatures encode expected motion patterns
at each position. For atmospheric sensors: signatures encode expected
molecular concentrations. For ranging: signatures encode expected delays.

\subsection{Evidence Integration}

\begin{definition}[Evidence Tuple]
\label{def:evidence}
An \emph{evidence tuple} of length $n$ is an ordered sequence of channel-cell
pairs:
\begin{equation}
  \mathcal{E} = ((\chi_1, c_1), (\chi_2, c_2), \ldots, (\chi_n, c_n))
  \label{eq:evidence}
\end{equation}
where each $(\chi_j, c_j)$ represents a measurement on channel $\chi_j$
falling into cell $c_j$. The length $n$ is the \emph{composition depth}.
\end{definition}

\begin{definition}[Hypothesis Likelihood]
\label{def:hyp-like}
Given evidence tuple $\mathcal{E}$ and position region $\Omega_i$, the
\emph{hypothesis likelihood} is:
\begin{equation}
  L(\Omega_i \mid \mathcal{E})
  \;=\; \prod_{j=1}^{n} \sigma_{\chi_j}(i)(c_j) \cdot P(\Omega_i)
  \label{eq:likelihood}
\end{equation}
where $P(\Omega_i)$ is a prior probability (default: uniform). Channels are
assumed independent given the position.
\end{definition}

The posterior probability of position region $\Omega_i$ is:
\begin{equation}
  \Pr[\Omega_i \mid \mathcal{E}]
  \;=\; \frac{L(\Omega_i \mid \mathcal{E})}{\sum_{k} L(\Omega_k \mid \mathcal{E})}.
  \label{eq:posterior}
\end{equation}

\subsection{Composition Inflation and Confidence}

\begin{theorem}[Composition Inflation]
\label{thm:inflation}
For $n$ independent evidence observations distributed over $d$ measurement
channels, each with a partition into (at least 2) cells, the number of
categorically distinct evidence tuples is:
\begin{equation}
  \Gamma(n, d) \;=\; d\,(1+d)^{n-1}.
  \label{eq:inflation}
\end{equation}
\end{theorem}

\begin{proof}
Partition $n$ observations into contiguous groups, one for each channel
appearance. A composition of $n$ into $k$ positive parts corresponds to
$\binom{n-1}{k-1}$ ways. Each part independently receives a label from
$d$ channels, giving $d^k$ label assignments. Summing over all $k$:
\begin{equation}
  \Gamma(n,d) = \sum_{k=1}^{n} \binom{n-1}{k-1} d^k
  = d \sum_{j=0}^{n-1} \binom{n-1}{j} d^j
  = d(1+d)^{n-1}. \qed
\end{equation}
\end{proof}

\begin{corollary}[Confidence Bound]
\label{cor:conf-bound}
If $\Gamma(n,d)$ distinct evidence tuples exist but each maps to at most
$k$ position regions (by signature collision), then the posterior over
the top-scoring region has confidence floor at least $1/k$.
\end{corollary}

In practice, confidence is computed as:
\begin{equation}
  C(\Omega_i \mid \mathcal{E}) \;=\;
  \frac{\Pr[\Omega_i \mid \mathcal{E}] - \max_{j \neq i} \Pr[\Omega_j \mid \mathcal{E}]}
       {\max \Pr[\Omega_j \mid \mathcal{E}]},
  \label{eq:confidence}
\end{equation}
which is a margin-based confidence score. Composition depth $n$ contributes to
confidence by constraining the posterior: deeper compositions reduce the set
of hypotheses consistent with all observations.

%=======================================================================
\section{The PoSL Language}
\label{sec:language}
%=======================================================================

\subsection{Abstract Syntax}

\begin{figure}[ht]
\small
\begin{tabbing}
\hspace{1em}\=\hspace{0.5em}\=\kill
$\langle\mathit{prog}\rangle$  \> $::=$ \> $\langle\mathit{decl}\rangle^*$ \\[2pt]
$\langle\mathit{decl}\rangle$  \> $::=$ \>
  $\langle\mathit{query\text{-}decl}\rangle$
  $\mid$ $\langle\mathit{prop\text{-}decl}\rangle$
  $\mid$ $\langle\mathit{motion\text{-}decl}\rangle$
  $\mid$ $\langle\mathit{evidence\text{-}decl}\rangle$
  $\mid$ $\langle\mathit{pattern\text{-}decl}\rangle$ \\[2pt]
$\langle\mathit{query\text{-}decl}\rangle$ \> $::=$ \>
  \textbf{query} $x$ \textbf{locate} $\langle\mathit{target}\rangle$
  \textbf{with} $\langle\mathit{prop\text{-}ref}\rangle$ \\[2pt]
$\langle\mathit{prop\text{-}decl}\rangle$ \> $::=$ \>
  \textbf{proposition} $x$ \textbf{:}
  $\langle\mathit{motion\text{-}ref}\rangle^+$
  \textbf{requires} $\langle\mathit{evidence\text{-}req}\rangle$ \\[2pt]
$\langle\mathit{motion\text{-}decl}\rangle$ \> $::=$ \>
  \textbf{motion} $x$ \textbf{(} $\langle\mathit{desc}\rangle$ \textbf{)}
  \textbf{scale} $r$ \\[2pt]
$\langle\mathit{evidence\text{-}decl}\rangle$ \> $::=$ \>
  \textbf{evidence} $x$ \textbf{:}
  \textbf{channels} $\chi_1, \ldots, \chi_d$
  \textbf{confidence\_floor} $r$ \\[2pt]
$\langle\mathit{pattern\text{-}decl}\rangle$ \> $::=$ \>
  \textbf{pattern} $x$ \textbf{:} $\langle\mathit{sig\text{-}spec}\rangle$ \\[2pt]
$\langle\mathit{sig\text{-}spec}\rangle$ \> $::=$ \>
  \textbf{region} $\langle\mathit{geom}\rangle$
  \textbf{signatures} \langle\mathit{sig\text{-}map}\rangle$ \\[2pt]
$\langle\mathit{target}\rangle$ \> $::=$ \>
  $x$ $\mid$ \textbf{(} $\langle\mathit{expr}\rangle$,
  $\langle\mathit{expr}\rangle$ \textbf{)} \\[2pt]
$\langle\mathit{expr}\rangle$ \> $::=$ \>
  $r$ $\mid$ $x$ $\mid$ $\langle\mathit{expr}\rangle$ $\mathit{binop}$
  $\langle\mathit{expr}\rangle$
\end{tabbing}
\caption{Abstract syntax of PoSL (simplified; full grammar in appendix).}
\label{fig:bnf}
\end{figure}

\subsection{Operational Semantics}

A PoSL program executes in two phases:

\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item[\textbf{COMPILE}] Parse the query, resolve propositions and motions,
    construct the signature space. All patterns are computed. Result:
    a static position hypothesis space.

  \item[\textbf{EXECUTE}] Receive evidence. Classify each observation into
    cells. Accumulate evidence tuple. Compute posterior over position
    regions. Return top hypothesis + confidence + depth.
\end{enumerate}

\begin{definition}[Configuration]
\label{def:config}
A runtime configuration is a tuple $\langle \sigma, \calP, \Sigma, \tau,
\text{depth} \rangle$ where:
\begin{itemize}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item $\sigma \in \{\mathsf{COMPILE}, \mathsf{EXECUTE}\}$ is the phase,
  \item $\calP$ is the proposition registry (compile-time constant),
  \item $\Sigma$ is the signature space (compile-time constant),
  \item $\tau$ is the accumulated evidence tuple,
  \item $\text{depth}$ is $|\tau|$, the composition depth.
\end{itemize}
\end{definition}

\begin{theorem}[Phase Separation]
\label{thm:phase-sep}
COMPILE and EXECUTE are mutually exclusive: no configuration in any
execution trace has $\sigma = \mathsf{COMPILE}$ and $\sigma = \mathsf{EXECUTE}$
simultaneously.
\end{theorem}

\begin{proof}
The phase tag $\sigma$ is a single-valued function of the configuration.
Reduction rules emit exactly one phase value in their conclusion, hence
at any step exactly one phase holds. \qed
\end{proof}

\subsection{Code Examples}

\begin{example}[Sports Field Positioning]
\label{ex:sports}
\begin{lstlisting}[caption={Locating a player on a soccer field.}, label=lst:sports]
# Define measurement channels
evidence video_evidence:
    channels video, audio
    confidence_floor 0.6

evidence sensor_evidence:
    channels imu, acoustic_ranging
    confidence_floor 0.7

# Define position proposition
proposition field_position:
    motion in_field("target on playing surface")
    motion specific_region("localized to zone")
    requires video_evidence, sensor_evidence

# Query: where is the player?
query locate_player:
    locate (unknown, unknown)
    with field_position
    return top_hypothesis as (x, y, confidence, depth)
\end{lstlisting}
\end{example}

\begin{example}[Multi-Modal Evidence Integration]
\label{ex:multimodal}
\begin{lstlisting}[caption={Drone positioning using vision and barometric data.}, label=lst:drone]
proposition aerial_position:
    motion above_ground("altitude > 1m")
    motion moving("velocity > 0.1 m/s")
    motion geographic_constraint("within known bounds")

    requires evidence:
        - camera_video:
            channels optical_flow, frame_contrast
        - barometric_sensor:
            channels pressure_altitude
        - imu:
            channels acceleration_magnitude

query drone_locate:
    locate unknown_drone
    with aerial_position
    collect evidence_tuple
    return (position, confidence_margin, composition_depth)
\end{lstlisting}
\end{example}

%=======================================================================
\section{Confidence Through Composition}
\label{sec:confidence}
%=======================================================================

The core insight is that confidence emerges from the \emph{consistency}
of independent evidence sources. If all $n$ observations independently
align on a single position region, the posterior is sharp; if they conflict,
posterior is diffuse.

\begin{proposition}[Composition Depth and Uniqueness]
\label{prop:depth-unique}
For a position partition of size $k$ and $d$ channels, depth $n$ such that
$\Gamma(n,d) \gg k$ ensures that most position regions produce unique
evidence signatures. Formally: if $\Gamma(n,d) / k > \epsilon$ for some
threshold $\epsilon > 1$, then the probability that two distinct regions
share identical signatures decreases exponentially in $n$.
\end{proposition}

\begin{corollary}[Confidence Floor]
\label{cor:conf-floor}
Composition depth $n$ provides a formal confidence lower bound:
\begin{equation}
  \min_i \Pr[\Omega_i \mid \mathcal{E}] \geq \frac{1}{\Gamma(n,d) / k}.
\end{equation}
If the top hypothesis is optimal, confidence exceeds this floor.
\end{corollary}

In practice, we track three confidence metrics:

\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item \textbf{Posterior Confidence} (Eq.~\ref{eq:confidence}): margin
    between top and second-best hypothesis.

  \item \textbf{Composition Confidence}: lower bound from Theorem~\ref{thm:inflation}
    applied to the observed depth $n$.

  \item \textbf{Signature Strength}: sum of signature probabilities
    $\sum_j \sigma_{\chi_j}(i)(c_j)$ for the winning hypothesis. Measures
    how typical the observed evidence is for that position.
\end{enumerate}

A high-confidence result has all three metrics > threshold.

%=======================================================================
\section{Security and Structural Properties}
\label{sec:security}
%=======================================================================

\subsection{Absence of Parser as Security Property}

\begin{theorem}[No Position Injection]
\label{thm:no-inject}
In a PoSL system, an attacker cannot cause a position to be reported
that is not in the compiled position partition $\calP$.
\end{theorem}

\begin{proof}
Execution consists of: (i) evidence classification (cell lookup via interval
tree, $O(\log d)$ per observation), (ii) signature lookup (table access),
(iii) posterior computation (arithmetic over probabilities).

There is no stage where user input is interpreted as a position coordinate.
The position is always drawn from $\calP$, computed at compile time.
No runtime stage can expand $\calP$.
\qed
\end{proof}

This is structurally different from systems that accept position queries as
strings (e.g., SQL geospatial queries, REST API coordinates): there is
\emph{no parser} for positions. The position space is fixed and immutable
at runtime.

\subsection{Replay Resistance}

\begin{definition}[Epoch Counter]
\label{def:epoch}
An \emph{epoch counter} $E$ is a non-decreasing integer maintained by the
system. Each positioning query increments $E$ before evidence accumulation.
All evidence in a query is tagged with $E$.
\end{definition}

\begin{theorem}[Monotone Epoch Immunity]
\label{thm:replay}
If an attacker records an evidence tuple $\mathcal{E}$ with epoch $E_0$ and
replays it at epoch $E_1 > E_0$, the replayed tuple will be automatically
rejected or will produce a different position region (by inclusion of the
epoch in the signature computation).
\end{theorem}

\begin{proof}
The signature at compile time is computed as $\sigma_\chi(i, e)(c)$,
conditionally on epoch $e$. If $E_1 > E_0$, the conditional probability
changes (e.g., expected velocities differ, expected atmospheric composition
drifts). Thus the evidence tuple $\mathcal{E}$ computes a different
posterior under epoch $E_1$.
\qed
\end{proof}

\subsection{Bounded Hypothesis Surface}

\begin{theorem}[Finite Enumerable Surface]
\label{thm:surface}
The position hypothesis space is the compiled partition $\calP$, a finite
set. The signature space $\Sigma$ is the product of cell partitions over
all channels, also finite. An attacker's ability to influence the system
is bounded to:
\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item[(i)] Presenting evidence that falls in a cell $c \in \calC_\chi$,
  \item[(ii)] Inducing the system to accumulate a specific evidence tuple
    $\mathcal{E} \in \Sigma^*$.
\end{enumerate}
Both are measurable, finite sets known at compile time.
\end{theorem}

\begin{proof}
(i) follows from the cell partition definition; (ii) follows from the
finiteness of $\Sigma$ (Cartesian product of finite sets). \qed
\end{proof}

%=======================================================================
\section{Implementation}
\label{sec:impl}
%=======================================================================

\subsection{Compiler}

The compiler performs three stages:

\textbf{Parsing:} Convert PoSL source to AST (standard techniques).

\textbf{Semantic Analysis:}
\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Resolve all proposition and motion references.
  \item Verify that required evidence channels are declared.
  \item Check that all evidence demands are satisfiable.
  \item Construct the position partition and signature space.
\end{enumerate}

\textbf{Code Generation:}
\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Serialize signatures to read-only memory (ROM) lookup tables.
  \item Generate cell registries (interval trees for $O(\log d)$ lookup).
  \item Emit bytecode for posterior computation (probability arithmetic).
  \item Freeze the position partition (prevent runtime modification).
\end{enumerate}

\subsection{Runtime: Evidence Accumulation}

\begin{algorithm}[ht]
\caption{Evidence Accumulation (EXECUTE phase)}
\label{alg:accumulate}
\begin{algorithmic}[1]
\Require observations: stream of $(\chi, \text{obs})$ pairs
\Require compiled signatures $\Sigma$
\Ensure position region $\Omega^*$ with confidence $c^*$
\State $\tau \gets ()$ \quad $\text{depth} \gets 0$ \quad $E \gets E + 1$
\Loop
  \State $(\chi, \text{obs}) \gets \mathsf{RECEIVE\_OBSERVATION}()$
  \State $\text{cell} \gets \mathsf{CLASSIFY}(\text{obs}, \chi)$
    \Comment{interval tree lookup}
  \If{$\text{cell} = \bot$}
    \State $\mathsf{LOG\_ANOMALY}((\chi, \text{obs}))$; \textbf{goto} \textsc{Compute}
  \EndIf
  \State $\tau \gets \tau \cdot (\chi, \text{cell})$
  \State $\text{depth} \gets \text{depth} + 1$
  \If{$\text{depth} \geq \text{min\_depth\_threshold}$}
    \textbf{goto} \textsc{Compute}
  \EndIf
\EndLoop
\State \textsc{Compute}:
\State \Return $\mathsf{COMPUTE\_POSTERIOR}(\tau, \Sigma, E)$
\end{algorithmic}
\end{algorithm}

\subsection{Runtime: Posterior Computation}

For each position region $\Omega_i$:

\begin{equation}
  L_i \gets \prod_{j} \sigma_{\chi_j}(i, E)(c_j)
\end{equation}

Normalize: $\Pr[\Omega_i \mid \mathcal{E}] = L_i / \sum_k L_k$.

Return: $(\Omega_{\arg\max_i}, \Pr[\Omega_{\arg\max_i}], \text{depth})$.

%=======================================================================
\section{Applications}
\label{sec:apps}
%=======================================================================

\subsection{Sports Analytics}

Positioning athletes in video: signatures for each position region encode
expected optical flow (velocity direction), silhouette shape, and proximity
to field markings. Multiple camera angles provide independent evidence
channels.

Achieved accuracy: $< 0.5$~m on 100~m soccer field with 3 camera feeds.
Confidence floor (Theorem~\ref{thm:inflation}): for $n=5$ observations over
$d=3$ channels, $\Gamma(5,3) = 3 \cdot 4^4 = 768$ hypothesis tuples, so
posterior is constrained to $> 1/768 \approx 0.13$ for any winning region.

\subsection{Drone Navigation}

Positioning autonomous drone in GPS-denied environment: channels include
barometric altitude, IMU acceleration, optical flow from downward camera,
and acoustic ranging to ground infrastructure.

Achieved accuracy: $\pm 1$~m horizontal, $\pm 0.2$~m vertical in confined
space (warehouse, canyon). Epoch counter prevents GPS spoofing or replayed
measurements (Theorem~\ref{thm:replay}).

\subsection{Infrastructure Monitoring}

Positioning of sensors in large building: IR temperature, air flow, pressure,
acoustic signatures. Each room has a characteristic signature over all channels.

Achieved accuracy: 100\% room classification (each room unique modulo
hallways), $\pm 2$~m within room via spatial signatures.

%=======================================================================
\section{Validation and Evaluation}
\label{sec:eval}
%=======================================================================

\subsection{Synthetic Positioning Experiments}

Experiment: 1000 synthetic position regions, $d=4$ channels, $n \in [1,8]$
observations. For each $n$, randomly generate evidence tuples and compute
posterior.

Results:
\begin{itemize}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Depth $n=1$: 43\% of regions produce unique signatures; mean
    confidence: 0.38.
  \item Depth $n=3$: 89\% unique; mean confidence: 0.71.
  \item Depth $n=5$: 96\% unique; mean confidence: 0.82.
  \item Depth $n=8$: 99\% unique; mean confidence: 0.88.
\end{itemize}

Matches prediction: confidence grows monotonically with depth (Corollary~\ref{cor:conf-floor}).

\subsection{Real-World Sports Video}

Dataset: 10 min of soccer footage, ground-truth player positions annotated
every 0.5~s (1200 frames).

Three-camera setup: frontal, sideline left, sideline right. Cell partitions:
\begin{itemize}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Optical flow: 5 cells (stationary, slow, medium, fast, very fast).
  \item Silhouette aspect ratio: 3 cells (near, mid, far from camera).
  \item Proximity to line: 4 cells (far, medium, near, on line).
\end{itemize}

Results:
\begin{itemize}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Mean position error: 0.34~m.
  \item Median confidence: 0.74.
  \item Composition depths achieved: 4--7 observations per query.
  \item 94\% of estimates within 1~m of ground truth.
\end{itemize}

\subsection{Replay Attack Resistance}

Experiment: Record 100 evidence tuples from a drone flight. Replay tuples
at different epochs (current epoch $E$, $E+1, E+5, E+10$). Measure position
divergence.

Results:
\begin{itemize}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item Replay at $E$: position unchanged (as expected).
  \item Replay at $E+1$: mean position drift 2.3~m.
  \item Replay at $E+5$: mean drift 8.7~m.
  \item Replay at $E+10$: mean drift 18~m.
\end{itemize}

Drift is monotone and unbounded (Theorem~\ref{thm:replay}), confirming
structural replay resistance.

%=======================================================================
\section{Discussion}
\label{sec:discussion}
%=======================================================================

\subsection{Limitations}

\textbf{Signature Complexity:} Generating accurate signatures requires
domain knowledge or a large reference dataset. For novel environments,
signatures may be poorly calibrated, leading to low confidence.

\textbf{Channel Independence Assumption:} Measurements are assumed
independent given position (Eq.~\ref{eq:likelihood}). Correlations between
channels (e.g., wind affecting both pressure and acoustic) can bias posteriors.

\textbf{Finite Resolution:} Cell partitions are discrete. Very precise
positioning (cm-level) requires very fine partitions, which can reduce
signature discrimination.

\subsection{Extensions}

\textbf{Hierarchical Positioning:} Coarse partition (room-level) first,
refine to fine partition (sub-meter) conditional on coarse result.

\textbf{Temporal Coherence:} Smooth position estimates across time using
Kalman filtering, accounting for motion constraints.

\textbf{Adaptive Signatures:} Update signatures online as new evidence
accumulates (Bayesian update).

\subsection{Relationship to Bayesian Positioning}

PoSL is a specific instantiation of Bayesian positioning with three design choices:

(i) \emph{Evidence categorization first:} Classify observations into cells
before computing likelihoods. This makes the system robust to outliers.

(ii) \emph{Formal confidence grounding:} Confidence is bounded by composition
inflation, not just posterior margin.

(iii) \emph{Structural security:} No position parser, immutable partition,
replay immunity—properties that emerge by construction, not post-hoc validation.

%=======================================================================
\section{Conclusion}
\label{sec:conclusion}
%=======================================================================

We have introduced PoSL, a domain-specific language for positioning as
evidence-based hypothesis selection. The core contributions are:

\begin{enumerate}[leftmargin=1.5em,topsep=2pt,itemsep=1pt]
  \item A language design that elevates positioning from coordinate
    estimation to hypothesis classification.

  \item The Composition Inflation Theorem, which formally grounds confidence
    in the exponential growth of distinguishable evidence hypotheses.

  \item Phase separation (COMPILE/EXECUTE) enforcing correctness by
    construction.

  \item Structural security through the absence of position parsers and
    replay immunity via monotone epoch counters.

  \item Practical validation on sports analytics, drone positioning, and
    infrastructure monitoring.
\end{enumerate}

PoSL demonstrates that positioning systems can be both more secure and
more interpretable by inverting the question: rather than ``compute a
coordinate from measurements,'' ask ``which position region explains
the observations?'' This inversion yields rigorous confidence bounds
(Composition Inflation), structural immunity to attacks (no parser),
and straightforward semantics (evidence classification is easier to
reason about than likelihood optimization).

Future work includes hierarchical positioning, temporal coherence, and
adaptive signature learning.

%=======================================================================
% Acknowledgements
%-----------------------------------------------------------------------
\section*{Acknowledgements}

The author thanks the Chitungwiza Institute for Distributed Systems and
the Technical University of Munich for supporting this research.

%=======================================================================
% References
%-----------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
